New lower bounds for Schur and weak Schur numbers

### 2021-04-09 article/article.tex
explication intuitive + valuers

\par
The fifth section gives new insight on the method that was formerly used to achieve new lower bounds. The principle behind it
might have discarded many interesting partititions of the search space and lowered the highest value that can be reached within it.
However, algorithms based on randomness such as Monte-Carlo algorithms may appear very useful if used in a search space with more
potential. For example, such an algorithm could be used to find better templates and improve inequalities and lower bounds for high values
of \(n\). This could be the subject fo a future work. Nevertheless, it would not help for the smallest values.

@@

\end{tabular}
\end{center}
\section{WSF-templates}
\begin{center}
WSF-template partitionning \([\![1, 42]\!]\) into 4 subsets
\begin{tabular}{|*{2}{c|}}
	\hline
	1 & 1, 2, 4, 8, 11, 22, 25 & \textbf{(66)}\\
	\hline
	2 & 5, 6, 7, 19, 21, 23, 36\\
	\hline
	3 & 9, 10, 12, 13, 14, 15, 16, 17, 18, 20\\
	\hline
	4 & 24, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 37, 38, 39, 40\\
	& 41, 42\\
	\hline
\end{tabular}
\end{center}
This template produces the inequality \(WS(n+3) \geqslant 42S(n) + 24\)
by placing 66 in the first subset.

>>>>

\par
The fifth section gives new insight on the method that was formerly used to achieve new lower bounds. The principle behind it
might have discarded many interesting partitions of the search space and lowered the highest value that can be reached within it.
However, algorithms based on randomness such as Monte-Carlo algorithms may appear very useful if used in a search space with more
potential. For example, such an algorithm could be used to find better templates and improve inequalities and lower bounds for high values
of \(n\). This could be the subject fo a future work. Nevertheless, it would not help for the smallest values.

@@

\end{tabular}
\end{center}
\section{WSF-templates}
\begin{center}
WSF-template partitionning \([\![1, 42]\!]\) into 4 subsets
\begin{tabular}{|*{2}{c|}}
	\hline
	1 & 1, 2, 4, 8, 11, 22, 25\\
	\hline
	2 & 5, 6, 7, 19, 21, 23, 36\\
	\hline
	3 & 9, 10, 12, 13, 14, 15, 16, 17, 18, 20\\
	\hline
	4 & 24, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 37, 38, 39, 40\\
	& 41, 42\\
	\hline
\end{tabular}
\end{center}

### 2021-04-09 article/weakschur.tex
explication intuitive + valuers

\begin{theorem}
Let \((p,q), (n,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \([\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]\) into \(n+k\)
weakly sum-free subsets.
\end{theorem}

@@

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } WS(n+k) \geqslant S(k) \left (WS(n) + \left \lceil \frac{WS(n)}{2}
\right \rceil +1 \right) + WS(n)\)
\end{corollary}

@@

\renewcommand{\arraystretch}{2}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\hline
  & & & & & \cellcolor{green}1 & \cellcolor{green}2 & \cellcolor{blue} 3&\cellcolor{green}4&5&6&7&8 \\
\hline
   \cellcolor{red}9 & \cellcolor{red}10 & \cellcolor{red}11 & \cellcolor{red}12 &\cellcolor{red}13&\cellcolor{red}14&\cellcolor{red}15&\cellcolor{red}16&\cellcolor{red}17&18&19&20&21\\
\hline
\cellcolor{yellow}22&\cellcolor{yellow}23&\cellcolor{yellow}24&\cellcolor{yellow}25&\cellcolor{yellow}26&\cellcolor{yellow}27&\cellcolor{yellow}28&\cellcolor{yellow}29&\cellcolor{yellow}30&31&32&33&34\\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
...&...&...&...&...&...&...&...&...&...&...&...&...\\
\hline
\end{tabular}

>>>>

\begin{theorem}
Let \((p,q), (n,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \([\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]\) into \(n+k\)
weakly sum-free subsets.
\end{theorem}

@@

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } WS(n+k) \geqslant S(k) \left (WS(n) + \left \lceil \frac{WS(n)}{2}
\right \rceil +1 \right) + WS(n)\)
\end{corollary}

\begin{remark}
In the above ineqality, a "\(+1\)" can be added to the lower bound if \(WS(n)\) is odd (more generally if \(q\) is odd in the theorem). However, it makes the proof less 
clear and it is never useful in practice.
\end{remark}

We first give an intuitive explanation of the above theorem. Let \((p, q) \in \mathbb{N}^2\) such that there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets. Let \(a \in \mathbb{N}\) with \(a > q\) and let's try to build a colouring of \([\![1, ap + q]\!]\) 
into \(n + k\) weakly sum-free subsets. Let \(l = a - b - 1\), \(r \in [\![1,q]\!]\) and \(w = a - l - r - 1 = b - r\). \\

First, put the integers of \([\![1, ap + q]\!]\) in the following table (with \(a\) columns and \(p + 1\) lines) and divide it into 3 blocks (the columns are numbered from \(-l\) to \(+q\)):

\begin{itemize}
	\item \(\mathcal{T}\) (the "tail"): the integers from 1 to \(q\). NB: this is line number 0.
	\item \(\mathcal{L}\) (the "lines"): the integers in columns \(-l\) to \(+r\) (excluding  \(\mathcal{T}\)).
	\item \(\mathcal{C}\) (the "columns"): the integers in the last \(w\) columns (excluding  \(\mathcal{T}\)).
\end{itemize}

Note that with this numbering of columns, the column of the sum of two numbers is the only integer in \([\![-l,q]\!]\) equal to two the sum of the columns modulo \(a\).

\renewcommand{\arraystretch}{2}
\begin{tabular}{|c|c|c|c|c|c|c|c|c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{blue}}c|>{\columncolor{green}} c|}
\hline
  & & & & & \cellcolor{green}1 & \cellcolor{green}2 & \cellcolor{blue} ...&\cellcolor{green}\(r\)&\(r+1\)&...&\(b - 1\)&\(b\) \\
\hline
   \cellcolor{red}9 & \cellcolor{red}10 & \cellcolor{red}11 & \cellcolor{red}12 &\cellcolor{red}13&\cellcolor{red}14&\cellcolor{red}15&\cellcolor{red}16&\cellcolor{red}17&18&19&20&21\\
\hline
\cellcolor{yellow}22&\cellcolor{yellow}23&\cellcolor{yellow}24&\cellcolor{yellow}25&\cellcolor{yellow}26&\cellcolor{yellow}27&\cellcolor{yellow}28&\cellcolor{yellow}29&\cellcolor{yellow}30&31&32&33&34\\
\hline
...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...\cellcolor{yellow}&...&...&...&...\\
\hline
\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&\cellcolor{red}...&...&...&...&...\\
\hline
...&...&...&...&...&...&...&...&...&...&...&...&...\\
\hline
\end{tabular}

\underline{\textbf{\(\mathcal{T}\) block}} \\
We color this block using the weakly sum-free colouring of \([\![1,q]\!]\) with colors \(1, ..., n\).

\underline{\textbf{\(\mathcal{L}\) block}} \\
In this block,  we use the colors \(n + 1, ..., n + k\). We colour an integer \(x\) according to its line number (written \(\lambda(x)\)).
For every \(x \in \mathcal{L}\), we colour \(x\) with \(n + c\) where \(c\) is the colour of \(\lambda(x)\) in the sum-free colouring of  \([\![1,p]\!]\).
Let \((x, y) \in \mathcal(L)^2\). The cases are twofold.

\begin{itemize}
	\item \underline{\(\lambda(x+y) = \lambda(x) + \lambda(y)\)} \\
	In this case, we use the sum-free property of the colouring of \([\![1,p]\!]\) (in block \(\mathcal{C}\), we only use colours \(1, ..., n\)).

	\item \underline{\(\lambda(x+y) \neq \lambda(x) + \lambda(y)\)} \\
	In this case, we do not have information about the colour of \(\lambda(x+y)\). Thereby, we want to have \(x+y \in \mathcal{C}\).
	A simple solution is to limit the horizontal movement, that is if the sum changes line, not to move too far so that it stays in \(\mathcal{C}\). 
	There, the maximal displacement to the left (resp. to the right) is \(2l\) (resp. \(2r\)). Not crossing entirely \(\mathcal{C}\) by going to the left 
	is then expressed as \(-2l > -a + r\). Likewise, not going to far to the right is expressed as \(2r < a - l\). It can then be written as
	 \(\text{max}(l, r) \leqslant w\).
\end{itemize}

\underline{\textbf{\(\mathcal{C}\) block}} \\
In this block,  we use colors \(1, ..., n\). We colour an integer \(x\) according to its column number (written \(\tilde{\pi}(x)\), it is linked to the 
projection on the first line, written \(\pi\), by the relation \(\tilde{\pi}(x) = \pi(x) - a\)). A simple solution is to colour \(x\) with the same colour 
as \(\tilde{\pi}(x)\) in the weakly sum-free colouring of \([\![1,q]\!]\). As long as \(2b \leqslant a + r\) (not going two far to the right) and there 
is no \(x \in \tilde{\pi}(\mathcal{C})\) such that \(2x \in \tilde{\pi}(\mathcal{C})\) (so that we do not have a sum in \(\mathcal{C}\) when 
taking two numbers in the same column), the colours \(1, ..., n\) are sum-free. \\

In particular, taking \(w = l = \lceil \frac{q}{2} \rceil\) and \(r = \lfloor \frac{q}{2} \rfloor\) works, thus obtaining the above theorem. \\

@@
